Estimating the magnetization distribution within rectangular rock samples

### 2016-06-23 manuscript/manuscript.tex
include tests with 1, 2 and 3 observation planes generate pdf with twocolunms
include the new figures

%\documentclass[draft,gc]{agutex}
\documentclass[galley,gc]{agutex}

@@

\usepackage{lineno}
\linenumbers*[1]
%  To add line numbers to lines with equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables
%
%
% DO NOT USE \psfrag or \subfigure commands.
%
%
%  Uncomment the following command to include .eps files
%  (comment out this line for draft format):
%\usepackage[dvips]{graphicx}
%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
 %\setkeys{Gin}{draft=false}
%
% Substitute one of the following for [dvips] above
% if you are using a different driver program and want to
% proof your illustrations on your machine:
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
% See how to enter figures and tables at the end of the article, after
% references.
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

@@

Over the last decades, scanning magnetic microscopy techniques have 
been increasingly used in paleomagnetism and rock magnetism. 
Different from standard paleomagnetic magnetometers, scanning 
magnetic microscopes produce high-resolution maps of the vertical 
component of the magnetic induction field (flux density) 
on a plane located over the sample. 
These high-resolution magnetic maps can be used for estimating the 
magnetization distribution within a rock sample by inversion.
Previous studies have estimated the magnetization distribution 
within rock samples by inverting the magnetic data measured on a 
single plane above the sample. Here we present a new spatial domain
method for inverting the magnetic induction measured on four 
planes around the sample in order to retrieve its internal 
magnetization distribution. We have presumed that the internal 
magnetization distribution of the sample varies along one of
its axes. Our method approximates the sample geometry
by an interpretation model composed of a one-dimensional 
array of juxtaposed rectangular prisms with uniform magnetization. 
The Cartesian components of the magnetization vector 
within each rectangular prism 
are the parameters to be estimated by solving a linear inverse problem. 
Our method automatically deals with the averaging of 
the measured magnetic data due to the finite size of the magnetic sensor,
preventing the application of a deconvolution before the inversion.
Tests with synthetic data show the performance of our method in 
retrieving complex magnetization distributions even in 
the presence of magnetization heterogeneities. 
We have also applied our method to invert experimentally measured 
magnetic data produced by a highly-magnetized synthetic sample 
that was manufactured in the laboratory.
The results show that, even in the presence of apparent position
noise, our method was able to retrieve the magnetization distribution
consistent with the anhysteric remanance magnetization
induced in the sample.

@@

Based on the estimation of the total (bulk) remanent magnetization (RM), 
the standard paleomagnetic techniques have been developed for more than 
40 years in studies of rock magnetism, magnetostratigraphic and 
paleogeographic reconstructions [e.g., \citeauthor{voo1993}, \citeyear{voo1993}]. 
The evolution of the paleomagnetic techniques is directly linked to 
the development of rock magnetometers and data processing methods.

Since its beginning in the 1950s, 
palemagnetism has seen at least three generation of rock magnetometers 
\citep{turner2015}: astatic systems, 
spinner magnetometers and superconducting magnetometers. 
Astatic system was used during the earliest paleomagnetic studies, 
such as \citet{blackett1952} and Pozzi e Thellier (1960) FALTA O BIBTEX 
and it was quickly replaced by spinner magnetometers (e.g., Nagata, 1961) FALTA O BIBTEX. 
Spinner magnetometers are most used devices in the paleomagnetic 
laboratories but they have been systematically replaced by 
superconducting magnetometers. Superconducting rock magnetometers 
have been developed since 70’s \citep{dodson1974, goree-fuller1976}, 
reaching notable sensitivity and quickness of measurements on the last two decade 
\citep{kirschvink2015}. 
This kind of rock magnetometer can measure both discrete samples 
(single sample mode) and long-core samples (pass-through mode), 
with a good spatial accuracy ($\sim$1 cm).
The long-core samples or u-channel samples are essential to 
investigations of magnetostratigraphy and environmental magnetism, 
such as those ones obtained at IODP (International Ocean Drilling Project).

@@

Notice that the forward problem presented in this subsection
describes the averaged magnetic induction produced by the 
interpretation model at the observation points. The averaging
is calculated over a regular grid of $Q$ points on the
active area $A$ of the magnetic sensor 
(Figure \ref{fig:b-mag-flux-approx}).

@@

We have applied our method to invert the synthetic data produced
by a simulated rock sample (not shown) with $L_{x} = 16$~mm and 
$L_{y} = L_{z} = 3$~mm, according to the Figure \ref{fig:sample}.
This synthetic sample is formed by $P = 16$ juxtaposed prisms along 
the $x$-axis, where each prism has a uniform and high magnetization
(red dots in Figures \ref{fig:estimate-validation},
\ref{fig:estimate-pre-processing} and \ref{fig:estimate-sensor2sample}).
In all tests presented in this section, 
the simulated magnetic sensor has an square active area  
$A = (300 \, \mu \text{m})^{2}$, which was discretized into
a regular grid of $7 \times 7$ points, totalling $Q = 49$
(Figure \ref{fig:b-mag-flux-approx}). Besides,
the magnetic data produced by the synthetic sample on 
each plane $\alpha = 0, 1, 2, 3$ were calculated at
$N = 4,284$ points, on a set of $42$ regularly spaced lines 
that are parallel to the $x$-axis, each one with $102$ regularly
spaced points. The data were also contaminated with a 
pseudo-random Gaussian noise having null mean and standard 
deviation equal to $30,000$~$\mu$T. These data simulate
an observed data set.
Besides, we have simulated the misalignment problems 
described in the subsection \ref{subsec:Pre-processing}.
Table \ref{tab:misalignment-parameters} shows the parameters
$\theta$, $\Delta x^{\prime}$ and $\Delta y^{\prime}$ (Figure 
\ref{fig:acquisition_errors}b) representing these misalignments
in the magnetic data produced by the synthetic sample on
the planes $\alpha = 0, 1, 2, 3$.

>>>>

%\documentclass[draft,gc]{agutex}
%\documentclass[galley,gc]{agutex}
\documentclass[twocolumn,gc]{agutex}

@@

%\usepackage{lineno}
%\linenumbers*[1]
%  To add line numbers to lines with equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables
%
%
% DO NOT USE \psfrag or \subfigure commands.
%
%
%  Uncomment the following command to include .eps files
%  (comment out this line for draft format):
%\usepackage[dvips]{graphicx}
%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
 %\setkeys{Gin}{draft=false}
%
% Substitute one of the following for [dvips] above
% if you are using a different driver program and want to
% proof your illustrations on your machine:
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
% See how to enter figures and tables at the end of the article, after
% references.
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

@@

Over the last decades, scanning magnetic microscopy techniques have 
been increasingly used in paleomagnetism and rock magnetism. 
Different from standard paleomagnetic magnetometers, scanning 
magnetic microscopes produce high-resolution maps of the vertical 
component of the magnetic induction field (flux density) 
on a plane located over the sample. 
These high-resolution magnetic maps can be used for estimating the 
magnetization distribution within a rock sample by inversion.
Previous studies have estimated the magnetization distribution 
within rock samples by inverting the magnetic data measured on a 
single plane above the sample. Here we present a new spatial domain
method for inverting the magnetic induction measured on four 
planes around the sample in order to retrieve its internal 
magnetization distribution. We have presumed that the internal 
magnetization distribution of the sample varies along one of
its axes. Our method approximates the sample geometry
by an interpretation model composed of a one-dimensional 
array of juxtaposed rectangular prisms with uniform magnetization. 
The Cartesian components of the magnetization vector 
within each rectangular prism 
are the parameters to be estimated by solving a linear inverse problem. 
Our method automatically deals with the averaging of 
the measured magnetic data due to the finite size of the magnetic sensor,
preventing the application of a deconvolution before the inversion.
Tests with synthetic data show the performance of our method in 
retrieving complex magnetization distributions even in 
the presence of magnetization heterogeneities. 
We have also applied our method to invert experimentally measured 
magnetic data produced by a highly-magnetized synthetic sample 
that was manufactured in the laboratory.
The results show that, even in the presence of apparent position
noise, our method was able to retrieve the magnetization distribution
consistent with the isothermal remanance magnetization
induced in the sample.

@@

Based on the estimation of the total (bulk) remanent magnetization (RM), 
the standard paleomagnetic techniques have been developed for more than 
40 years in studies of rock magnetism, magnetostratigraphy and 
paleogeographic reconstructions [e.g., \citeauthor{voo1993}, \citeyear{voo1993}]. 
The evolution of the paleomagnetic techniques is directly linked to 
the development of rock magnetometers and data processing methods.

Since its beginning in the 1950s, 
paleomagnetism has seen at least three generation of rock magnetometers 
\citep{turner2015}: astatic systems, 
spinner magnetometers and superconducting magnetometers. 
Astatic system was used during the earliest paleomagnetic studies, 
such as \citet{blackett1952} and \citet{pozzi-thellier1963},
and it was quickly replaced by spinner magnetometers \citep{nagata1961}. 
Spinner magnetometers are most used devices in the paleomagnetic 
laboratories but they have been systematically replaced by 
superconducting magnetometers. Superconducting rock magnetometers 
have been developed since 70’s \citep{dodson1974, goree-fuller1976}, 
reaching notable sensitivity and quickness of measurements on the last two decade 
\citep{kirschvink2015}. 
This kind of rock magnetometer can measure both discrete samples 
(single sample mode) and long-core samples (pass-through mode), 
with a good spatial accuracy ($\sim$1 cm).
The long-core samples or u-channel samples are essential to 
investigations of magnetostratigraphy and environmental magnetism, 
such as those ones obtained at IODP (International Ocean Drilling Project).

@@

Notice that the forward problem presented in this subsection
can be easily modified to the case in which the magnetic data
are measured on less than four observation planes.
Besides, this forward problem
describes the averaged magnetic induction produced by the 
interpretation model at the observation points. The averaging
is calculated over a regular grid of $Q$ points on the
active area $A$ of the magnetic sensor 
(Figure \ref{fig:b-mag-flux-approx}).

@@

We have applied our method to invert the synthetic data produced
by a simulated rock sample (not shown) with $L_{x} = 16$~mm and 
$L_{y} = L_{z} = 3$~mm, according to the Figure \ref{fig:sample}.
This synthetic sample is formed by $P = 16$ juxtaposed prisms along 
the $x$-axis, where each prism has a uniform and high magnetization
(red dots in Figures \ref{fig:estimate-validation},
\ref{fig:estimate-pre-processing} and \ref{fig:estimate-sensor2sample}).
In all tests presented in this section, 
the simulated magnetic sensor has an square active area  
$A = (300 \, \mu \text{m})^{2}$, which was discretized into
a regular grid of $7 \times 7$ points, totalling $Q = 49$
(Figure \ref{fig:b-mag-flux-approx}). Besides,
the magnetic data produced by the synthetic sample on 
each plane $\alpha = 0, 1, 2, 3$ were calculated at
$N = 4,284$ points, on a set of $42$ regularly spaced lines 
that are parallel to the $x$-axis, each one with $102$ regularly
spaced points. The data were also contaminated with a 
pseudo-random Gaussian noise having null mean and standard 
deviation equal to $30,000$~$\mu$T. These data simulate
an observed data set.
Besides, we have simulated the misalignment problems 
described in the subsection \ref{subsec:Pre-processing}.
Table \ref{tab:misalignment-parameters} shows the parameters
$\theta$, $\Delta x^{\prime}$ and $\Delta y^{\prime}$ (Figure 
\ref{fig:acquisition_errors}b) representing these misalignments
in the magnetic data produced by the synthetic sample on
the planes $\alpha = 0, 1, 2, 3$.
All inversions shown in this section were run
